---
title: "Lecture_Notes_9-11-2024"
author: "Grace Rade"
date: "2024-09-11"
output: html_document
---
## S3, Subsetting

#### S3

```{r}
## defining a new generic
shuffle = function(x) {
  UseMethod("shuffle")
}

## now we define a method
shuffle.factor = function(f) {
  factor( sample(as.character(f)), levels = sample(levels(f)) )
}

shuffle.integer = function(x) {
  sample(x)
}

shuffle.default = function(x) {
  stop("Class ", class(x), " is not supported by shuffle.", call. = FALSE)
}

## test it
shuffle( 1:10 )
shuffle( factor(c("A","B","C","A")) )
```

#### Matrices 

R supports the creation of 2d data structures of atomic vector types. Generally these are formed via a call to `matrix()`. They can support any type of object.

Matrices in R use column major ordering (data is stored in by column). We can populate a matrix by row, but the data is still stored by column. 

They are atomic vectors with a dimension attribute. Matrices (and arrays) are just atomic vectors with a dim attribute attached, they do not have a class attribute, but they do have an implicit matrix (and array) class. Arrays are just an n-dimensional extension of matrices and are defined by adding the appropriate dimension sizes.

#### Data Frames

A data frame is how R handles heterogeneous tabular data and is one of the most commonly used data structures in R. R stores data frames using a *list of equal length vectors* (atomic or generic).

```{r}
(df = data.frame(
  x = 1:3, 
  y = c("a", "b", "c"),
  z = c(TRUE)
))

```

Previous to R v4.0, the default behavior of data frames was to convert character data into factors. Sometimes this was useful, but mostly it wasn’t. This behavior is controlled via the `stringsAsFactors` argument to data.frame (and related functions like read.csv, read.table, etc.).

#### Subsetting

```{r}
## Exercise 1

x = c(56, 3, 17, 2, 4, 9, 6, 5, 19, 5, 2, 3, 5, 0, 13, 12, 6, 31, 10, 21, 8, 4, 1, 1, 2, 5, 16, 1, 3, 8, 1, 3, 4, 8, 5, 2, 8, 6, 18, 40, 10, 20, 1, 27, 2, 11, 14, 5, 7, 0, 3, 0, 7, 0, 8, 10, 10, 12, 8, 82, 21, 3, 34, 55, 18, 2, 9, 29, 1, 4, 7, 14, 7, 1, 2, 7, 4, 74, 5, 0, 3, 13, 2, 8, 1, 6, 13, 7, 1, 10, 5, 2, 4, 4, 14, 15, 4, 17, 1, 9)


x[seq(2, length(x), by = 3)] ## Select every third value starting at position 2 in x

x[seq(1, length(x), by = 2)-1] ## Remove all values with an odd index (e.g. 1, 3, etc.)

x[(!seq_along(x) %in% seq(1, length(x), by = 4)) & (x %% 2 ==1)] ## Remove every 4th value, but only if it is odd

```

