---
title: "Lecture_Notes_9-9-2024"
author: "Grace Rade"
date: "2024-09-09"
output: html_document
---

## Lists, Attributes, and S3

#### Generic Vectors (lists)

Lists can contain a heterogeneous collection of R objects. These objects can be objects of any type, and they won't get coerced. Lists can also contain other lists, meaning they don't have to be flat. 

  + can hold functions and other lists, all sorta of fun stuff
  
#### Nulls

`NULL` is a special value in R that represents nothing, always has a length zero and a type and mode of "NULL", and cannot have any attributes. 0-length length coercion is a special case of length coercion when one of the arguments has length 0.

#### Attributes

Attributes are pieces of metadata about objects in R. The most important attributes will have helper functions to get and set them. 

For non-common attributes, or if you don't want to use the helper function, you can use the `attr()` function. 

```{r}
(x = c(L=1,M=2,N=3))

## set an attribute
attr(x, "names") = 1

## see the attribute
attributes(x)
```

We often use attributes when working with factor objects. Factor objects are how R represents categorical data (e.g. a vairable wherre there is a discrete set of possible outcomes)

```{r}
(x = factor(c("Sunny", "Cloudy", "Rainy", "Cloudy", "Cloudy")))

```

