---
title: "Lecture_Notes_8-28"
author: "Grace Rade"
date: "2024-08-28"
output: html_document
---
## Logic and Types in R

Almost everything is a vector. A vector is a collection of related values/objects/etc. Atomic vectors are collections of vectors of the same type, generic vectors (lists) are collections of *any* R object. 

### Types of Vectors

  + complex vectors: imaginary numbers, a collection of doubles
  
  + raw: raw data values
  
Inspecting vectors with the `typeof()` and `mode()` functions to return the type and mode of the object. Mode is a slightly higher level hierarchy (i.e. returns numeric for integers and doubles). 

R is a dynamically typed language, meaning types change. It will automatically convert between most types without raising warnings or errors. 

  + if there are two different types in an atomic vector, R will coerce them to be compatible in a fixed order
  
    - character -> double -> integer -> logical
    
  + built in operators and functions will generally attempt to coerce values to an appropriate type for the given operation (ex. mathematical functions need numerics or logical operators)
  
    + 0 is FALSE, anything $\ne 0$ is TRUE
    
#### Missing Values

Missing values are represented by an NA. There are different NA types for different types of objects (ex. a missing value in a logical vector will be a type logical, for character it will be a character). We love how R accounts for missing data and then does something reasonable about it. 

```{r}
## Ex. 

typeof(NA) ## default NA is a logical
typeof(NA_character_) ## woo it's a character
typeof(c(NA,"")) ## character again, coerced to character
typeof(NA+1L) ## integer
typeof(NA_complex_) ## a number so complex we don't know what it is
typeof(NA_real_) ## ooohhh a double
```

NAs are sticky, so operations that encounter an NA will propagate it throughout (remember using `summarize()`). Lots of functions will allow us to remove NAs, keep that in mind for things like mean. 

NAs aren't always sticky though. It's helpful for logical vectors to consider them as an unknown value that could take on any of the possible values. 

```{r}
TRUE & NA ## the NA could be false, so we return an NA meaning unknown value
FALSE & NA ## returns false b/c false is always false so it doesn't matter what the NA is
TRUE | NA
FALSE | NA
```

Other missing values are `NaN` for "not a number", `Inf` and `-Inf` for positive and negative infinity. Follows math rules. It's a little weird but unlikely to actually encounter these. 

```{r}
## some predicate functions for Inf and NaN

is.finite(Inf)
is.infinite(-Inf)
is.nan(NaN)
```

Coersion can be weird. 

### Exercise 1

What type are these vectors? 

```{r}
c(1, NA+1L, "C") ## Guess: character NA
c(1L / 0, NA) ## Guess: NA
c(1:3, 5) ## Guess: integer
c(3L, NaN+1L) ## Guess: NaN
c(NA, TRUE) ## Guess: logical
```

```{r}
typeof(c(1, NA+1L, "C")) ## Answer: character (the NA is coerced to a character, so is the rest)
typeof(c(1L / 0, NA)) ## Answer: double (Inf is a double, and double wins over logical NA)
typeof(c(1:3, 5)) ## Answer: double (5 is a double b/c not specified as int)
typeof(c(3L, NaN+1L)) ## Answer: double (NaN + an int is a double)
typeof(c(NA, TRUE)) ## Answer: logical (yay correct, NA is logical by default)
```

What's the type conversion hierarchy: character -> double -> integer -> logical

### Control Flow

**Logical Operators:**

  + x & y: vectorized and
  
  + x | y: vectorized or
  
  + !x: not (vectorized w/ no non-vectorized alternative)
  
  + xor(x, y): exclusive or (vector)
  
  + x && y: non-vectorized and
  
  + x || y: non-vectorized or
  
Vectorized means that when we do an operation, we pay attention to the fact that the object is a vector and do element-wise comparisons. Non-vectorized operations expect a single value input and returns an error if there are multiple elements. We usually only use && and/or || to take advantage of short circuit evaluation. Vectorization applied to math as well. 

**Recycling Rules:** If the lengths of the vector do not match, then the shorter vector has its values recycled to match the length of the longer vector. The longer one has to be a multiple of the shorter one, if not it will get funky. 

**Comparison Operators**

  + x < y: less than
  
  + x > y: greater than
  
  + x <= y: less than or equal to
  
  + x >= y: greater than or equal to
  
  + x != y: not equal to
  
  + x == y: equal to
  
  + x %in% y: in/contains, vectorized over x (returned value will have the length of x regardless of the length of y)
  
Can use comparison operators with character values. < and > can be a little unexpected but also work on character vectors. Alphabetical order (lowercase ordered before uppercase). 

**Conditions**

R has if and else statements, but they are not vectorized. So they only check a single value. We can use the `any()` or `all()` functions to do a bit of a work-around when necessary. We also have else if to do a secondary if statement (mutually exlcusive options). 

```{r}
x = 3

if (x < 0) { ## r checks the if
  "x is negative"
} else if (x > 0) { ## the if is False, so r checks the else if
  "x is positive"
} else { ## the else if is True, r does not check the else
  "x is zero"
}

x = 0

if (x < 0) { ## r checks the if
  "x is negative"
} else if (x > 0) { ## the if is False, so r checks the else if
  "x is positive"
} else { ## the else if is Fale, r checks teh else
  "x is zero"
}
```


Conditionals return a value invisibly. If you want to see it, you either have to return a print statement or save it to a variable and call it. 


#### Exercise 2

What do you expect the output to be?

```{r}
f = function(x) {
  # Check small prime
  if (x > 10 || x < -10) {
    stop("Input too big")
  } else if (x %in% c(2, 3, 5, 7)) {
    cat("Input is prime!\n")
  } else if (x %% 2 == 0) {
    cat("Input is even!\n")
  } else if (x %% 2 == 1) {
    cat("Input is odd!\n")
  }
}
f(1) ## Guess: "Input is odd!" (correct)
f(3) ## Guess: "Input is prime!" (correct)
f(8) ## Guess: "Input is even!" (correct)
f(-1) ## Guess: "Input is odd!" (correct)
f(-3) ## Guess: "Input is odd!" (correct)
f(1:2) ## Guess: warning b/c length of vector, "Input is odd!" (Answer: error, coercion)
f("0") ## Guess: error message, data type (correct)
f("3") ## Guess: error message, data type (Answer: error, input too big)
f("zero") ## Guess: error message, data type (Answer: error, input too big)
```

