---
title: "Lecture_Notes_9-16-2024"
author: "Grace Rade"
date: "2024-09-16"
output: html_document
---

## Tidy Data and tidyr

We love the tidyverse! Helps us work to clean data. <3

Tidy data is a format that is designed to be most effective for cleaning data and modeling it. To take from Tolstoy, "all tidy datasets are alike; every untidy dataset is untidy in it's own way". 

"Ragged dataset": not all observations have the same amount of data (ex. Hot 100 Songs, not al songs are on the Hot 100 for the same amount of time)

#### Tiddbles: Data frames but modern

Data frames as a sturcute can have some annoying behaviors. Tibbles smooth out those irritating quirks. The core features of tibbles is to have a nicer printing method as well as being “surly” and “lazy”. 

Ex. Iris dataset doesn't fit in output (horizontally across), so you have to scroll down to see some of the vairables. It's fine but that can be a huge issue when the dfs have a ton or rows. We also don't see what the types of the variables are. 

```{r}
library(tibble)

iris ## as a df, doesn't show all the relevant info
as.tibble(iris) ## we see the first 10 rows, see the types of the variables, and shows what the additional columns are (gorg)
```

Tibbles are "surly" and "lazy"

  + Lazy meaning they preserve the data type if they can (type consistent). Subsetting a tibble will always result in a tibble. This is different in column subsetting from data frames
  
    - Doesn't do partial matching
    
    - Only vectors with length 1 will undergo length coercion / recycling - anything else throws an error
    
  + Surly as in they will tell you when they don't like what you want them to do
  
Tibbles are just specialized data frames, and will fall back to the base data frame methods when needed.

#### magrittr

Magrittr’s pipe is an infix operator that allows us to link two functions together in a way that is readable from left to right. 

```{r}
g(x=1, y=2) %>%  # %>%  is the operator, |> is the new pipe
  f(n=2)
```

Base R has a pipe (`|>`) that you can use without loading tidyverse/magrittr/dplyr. I don't like it though. 

#### dplyr

```{r}
library(dplyr)
library(nycflights13)

flights |> select(starts_with("dep"), starts_with("arr")) ## starts_with is cool

flights |> select(where(is.numeric)) ## where is also cool, like a logical subset

flights |> relocate(carrier, origin, dest) ## move these columns to the start (or end)

flights |> pull("year") |> head() ## oooh pull

## you can also use the .by argument in summarize
```

The .by (and by) arguments are used for per operation grouping while group_by() is intended for persistent grouping.

```{r}
flights |> 
  summarize(
    n = n(), 
    min_dep_delay = min(dep_delay, na.rm=TRUE), 
    max_dep_delay = max(dep_delay, na.rm=TRUE),
    .by = origin
  )
```

