---
title: "Reading_Notes_10-9-2024"
author: "Grace Rade"
date: "2024-10-01"
output: html_document
---
## Web APIs

#### The httr2 package

```{r}
library(httr2)
```

httr2 provides a pipeable API with an explicit request object that solves more problems felt by packages that wrap APIs (e.g. built-in rate-limiting, retries, OAuth, secure secrets, and more).

We start by creating a request: 

```{r}
req <- request(example_url()) ## each request begins wtih a url, this is a server of the package so the vignette will run
req

req |> req_dry_run()
```

The first line of the request contains three important pieces of information:

  + The HTTP method, which is a verb that tells the server what you want to do. Here’s its GET, the most common verb, indicating that we want to get a resource. Other verbs include POST, to create a new resource, PUT, to replace an existing resource, and DELETE, to delete a resource.

  + The path, which is the URL stripped of details that the server already knows, i.e. the protocol (http or https), the host (localhost), and the port (42853).

  + The version of the HTTP protocol. This is unimportant for our purposes because it’s handled at a lower level.

The following lines specify the HTTP headers, a series of name-value pairs separated by :. The headers in this request were automatically added by httr2, but you can override them or add your own with req_headers():

```{r}
req |>
  req_headers(
    Name = "Hadley",
    `Shoe-Size` = "11",
    Accept = "application/json"
  ) |>
  req_dry_run()

## headers are case sensitive

## add a body to the request
req |>
  req_body_json(list(x = 1, y = "a")) |>
  req_dry_run()
```

  + The method has changed from GET to POST. POST is the standard method for sending data to a website, and is automatically used whenever you add a body. Use req_method() to for a different method.

  + There are two new headers: Content-Type and Content-Length. They tell the server how to interpret the body — it’s encoded as JSON and is 15 bytes long.

  + We have a body, consisting of some JSON

Different servers want data encoded differently so httr2 provides a selection of common formats. For example, req_body_form() uses the encoding used when you submit a form from a web browser. And req_body_multipart() uses the multipart encoding which is particularly important when you need to send larger amounts of data or complete files:

```{r}
req |>
  req_body_form(x = "1", y = "a") |>
  req_dry_run()

req |>
  req_body_multipart(x = "1", y = "a") |>
  req_dry_run()
```

To actually perform a request and fetch the response back from the server, call req_perform():

```{r}
req <- request(example_url()) |> req_url_path("/json")
resp <- req |> req_perform()
resp

resp |> resp_raw() ## see what we actually recieved
```

An HTTP response has a very similar structure to an HTTP request. The first line gives the version of HTTP used, and a status code that’s optionally followed by a short description. Then we have the headers, followed by a blank line, followed by a body. The majority of responses will have a body, unlike requests. You can extract data from the response using the resp_() functions:

  + resp_status() returns the status code and resp_status_desc() returns the description:
  
  + You can extract all headers with resp_headers() or a specific header with resp_header()
  
  + You can extract the body in various forms using the resp_body_*() family of functions. Since this response returns JSON we can use resp_body_json():
  
```{r}
resp |> resp_status()
resp |> resp_status_desc()

resp |> resp_headers()
resp |> resp_header("Content-Length")

resp |> resp_body_json() |> str()
```

Responses with status codes 4xx and 5xx are HTTP errors. httr2 automatically turns these into R errors. 


A number of req_ functions don’t directly affect the HTTP request but instead control the overall process of submitting a request and handling the response. These include:

  + req_cache() sets up a cache so if repeated requests return the same results, you can avoid a trip to the server.

  + req_throttle() will automatically add a small delay before each request so you can avoid hammering a server with many requests.

  + req_retry() sets up a retry strategy so that if the request either fails or you get a transient HTTP error, it’ll automatically retry after a short delay.