---
title: "Reading_Notes_9-18-2024"
author: "Grace Rade"
date: "2024-09-14"
output: html_document
---
## Rectangling

Rectangling is the art and craft of taking a deeply nested list (often sourced from wild caught JSON or XML) and taming it into a tidy data set of rows and columns. There are three functions from tidyr that are particularly useful for rectangling:

  + `unnest_longer()` takes each element of a list-column and makes a new row.

  + `unnest_wider()` takes each element of a list-column and makes a new column.

  + `hoist()` is similar to `unnest_wider()` but only plucks out selected components, and can reach down multiple levels.
  
    - `hoist()` removes the named components from the user list-column, so you can think of it as moving components out of the inner list into the top-level data frame.
  
```{r}
library(tidyr)
library(dplyr)
library(repurrrsive)
```

#### GitHub Users

```{r}
users <- tibble(user = gh_users) ## import the data, each value is a list (no good to us)

users %>% unnest_wider(user) ## we can use unnest_wider() to make each value of the list into a column

users %>% hoist(user, 
  followers = "followers", 
  login = "login", 
  url = "html_url"
) ## or we can use hoist() to only get the few columns we want, syntax similar to purrr:pluck()
```

#### GitHub Repos

```{r}
repos <- tibble(repo = gh_repos) ## similar import format to users

## this time the repos re a list of repositories that belong to that user. These are observations, so they should become rows

repos <- repos %>% 
  unnest_longer(repo) ## now they are rows

repos %>% hoist(repo, 
  login = c("owner", "login"), 
  name = "name",
  homepage = "homepage",
  watchers = "watchers_count"
) ## now they are variables
```

