---
title: "Reading_Notes_10-16-2024"
author: "Grace Rade"
date: "2024-10-09"
output: html_document
---

## Building APIs

```{r}
library(plumber)
```

Plumber allows you to create a web API by merely decorating your existing R source code with roxygen2-like comments. These comments allow plumber to make your R functions available as API endpoints. You can use either #* as the prefix or #', but we recommend the former since #' will collide with roxygen2. You can visit this URL using a browser or a terminal to run your R function and get the results. 

#### Web APIs

The Hypertext Transfer Protocol (HTTP) is the dominant medium by which information is exchanged on the Internet. An Application Programming Interface (API) is a broad term that defines the rules that guide your interaction with some software.Plumber allows you to create APIs by merely decorating your existing R code with special annotations. Plumber translates the annotations you place on your functions into an HTTP API that can be called from other machines on your network. 

```{r, eval = F}
# plumber.R

#* Echo the parameter that was sent in
#* @param msg The message to echo back.
#* @get /echo
function(msg=""){
  list(msg = paste0("The message is: '", msg, "'"))
}

#* Plot out data from the iris dataset
#* @param spec If provided, filter the data to only this species (e.g. 'setosa')
#* @get /plot
#* @serializer png
function(spec){
  myData <- iris
  title <- "All Species"

  # Filter if the species was specified
  if (!missing(spec)){
    title <- paste0("Only the '", spec, "' Species")
    myData <- subset(iris, Species == spec)
  }

  plot(myData$Sepal.Length, myData$Petal.Length,
       main=title, xlab="Sepal Length", ylab="Petal Length")
}
```

Once you have plumber installed, you can use the pr() function to translate this R file into a Plumber API:

```{r, eval = F}
root <- pr("plumber.R")
root

## The pr object now encapsulates all the logic represented in your plumber.R file. The next step is to bring the API to life using the pr_run() method:

root %>% pr_run()
```

You should see a message about your API running on your computer on port 8000. The API will continue running in your R session until you press the Esc key.
