---
title: "Reading_Notes_9-16-2024"
author: "Grace Rade"
date: "2024-09-14"
output: html_document
---

## Dplyr and Tidy Data

#### Tidy Data

There are three requirements for tidy data:

  1. Each variable goes in a columns
  
  2. Each observation goes in a row
  
  3. Each value goes in a cell
  
These are inter-related and it's impossible to only satisfy two of the three. In practice, you only need to put each dataset into a tibble and each variable into a column. 

The advantages of tidy data are that dataset format is standardized across different sources and that it is really easy to wrok with in R. However, not all data you encounter is tidy. This is becuase not a lot of people are familiar with tidy format, and data is often organized to make entry easy or for a specific purpose. 

Sometimes you will have to pivot the dataset to get it into tidy format, meaning that you might have to fix a variable spread across columns or an observation spread across rows. You can pivot longer or pivot wider, in order to get the data tidy. 

```{r}
library(tidyverse)

## pivot longer: column names are not names of variables, but values of a variable

table4a ## the variable names 1999 and 2000 are the values of the year variable

table4a %>% 
  pivot_longer(c(`1999`, `2000`), names_to = "year", values_to = "cases") ## fixed it, now there are more rows
```

* note that the columns to pivot are chosen using `select()` notation

```{r}
## pivot wider: observations take up multiple rows

table2 ## there are two rows for each observation becuase cases and population aren't variables

table2 %>%
    pivot_wider(names_from = type, values_from = count) ## now there are more columns

```

There are also the `separate()` and `unite()` functions to deal with un-tidy data. `separeate()` pulls apart one column into multiple columns, by splitting wherever a separator character appears.

```{r}
table3 ## rate has cases/population as a character value, we want them separate

table3 %>% 
  separate(rate, into = c("cases", "population")) ## neat, the variables are still characters

## we can also specify the separator in the function and whether or not we want to convert the variables into differet types

table3 %>% 
  separate(rate, into = c("cases", "population"), sep ="/", convert = TRUE)

## we can also specify to separate after a certain number of characters

new <- table3 %>% 
  separate(year, into = c("century", "year"), sep = 2)

```

`unite()` puts multiple columns into a single columns. 

```{r}
new ## as above

table5 %>% 
  unite(new_year, century, year, sep = "") ## so below

```

