---
title: "Reading_Notes_8-28-2024"
author: "Grace Rade"
date: "2024-08-26"
output: html_document
---
## Vectors

Vectors have three properties:
  
  * type (use `typeof()` or `str()` to check)
  
  * length
  
  * attributes (use `attributes()` to check) aka arbitrary metadata
  
Atomic vectors must contain elements of all the same type, but lists can contain elements of multiple types. 

Atomic vectors include logical, integer, doubles, and character. They are always flat, even if you nest vectors, and created with `c()`. When combining vectors, if not all elements of both are the same type, the vectors are coerced to the most flexible type. Sometimes NAs are introduced by coercion.

  * integers: whole numbers (ints), a type of numeric vector
  
  * doubles: whole numbers of decimals, a numeric vector
    
    + `is.numeric()` returns TRUE for both doubles and integers as it checks the "number-liness" of the vector
  
  * characters: a vector of letters/strings or numbers encased in ""
  
  * logical: a vector containing TRUE/FALSE values (can be coerced into 1/0)
  
You can coerce vectors into different types, using an `as.type()` function, and it is often done automatically. 

Lists are different from atomic vectors as they can contain different types of objects, even lists. They are created using `list()`. They are sometimes called recursive vectors because of their ability to contain other lists. `c()` will combine several lists into one. 

Lists are also used to build up more complicated data structures, like data frames.

## Attributes

Attributes are arbitrary metadata about an object, stored as a list. They can be accessed individually with `attr()` or all together with `attributes()`. 

```{r}
y <- 1:10 ## create a vector
attr(y, "my_attribute") <- "This is a vector" ## assign an attribute
attr(y, "my_attribute") ## look there it is
str(attributes(y))
attributes(y) # here it is again
```

The structure function returns a new object with modified attributes. 

```{r}
structure(1:10, my_attribute = "This is a vector")
```

Generally, attributes are lost when an object is modified, but three are always kept:

  * Names: a character vector giving each element int he vector a name
    
    + you can name a vector when you create it, when you modify it, and when you copy it
    
      - `x <- c(a = 1, b = 2, c = 3)`
      
      - `x <- 1:3; names(x) <- c("a", "b", "c")`
      
      - `x <- setNames(1:3, c("a", "b", "c"))`
      
  
  * Dimensions: used to turn vectors into matrices and arrays
  
  * Class: used to implement the S3 OOP system
  
A factor is a vector that can contain only predefined values, and is used to store categorical data. Factors are built on top of integer vectors using two attributes: the class, `factor`, which makes them behave differently from regular integer vectors, and the `levels`, which defines the set of allowed values.

Factors are useful when you know the possible values an object might have, but not the specific values of each element in the object. 

```{r}
sex_char <- c("m", "m", "m") ## make a vector
sex_factor <- factor(sex_char, levels = c("m", "f")) ## turn it into a factor with the  levels m and f

table(sex_char)  ## just shows the data that is there

table(sex_factor) ## shows both categories, even though there are no f in the object

```

## Matrices and Arrays

Arrays are multi-dimensional atomic vectorswith a `dim` attribute. Matrices are special case arrays with two dimensions. THey can be created with `matrix()` or `array()`. 

```{r}
# Two scalar arguments to specify rows and columns
a <- matrix(1:6, ncol = 3, nrow = 2)
# One vector argument to describe all dimensions
b <- array(1:12, c(2, 3, 2))

# You can also modify an object in place by setting dim()
c <- 1:6
dim(c) <- c(3, 2)
c

dim(c) <- c(2, 3)
c

```

`length()` generalizes to `nrow()` and `ncol()` for matrices and `dim()` for arrays. `names()` generalizes to generalizes to `rownames()` and `colnames()` for matrices, and `dimnames()`, a list of character vectors, for arrays. 

```{r}
length(a) 
#> [1] 6
nrow(a)
#> [1] 2
ncol(a)
#> [1] 3
rownames(a) <- c("A", "B") ## set the row names
colnames(a) <- c("a", "b", "c") ## set the column names
a


length(b)
#> [1] 12
dim(b)
#> [1] 2 3 2
dimnames(b) <- list(c("one", "two"), c("a", "b", "c"), c("A", "B"))
b
```


`c()` generalizes to `cbind()` and `rbind()` for matrices and `abind()` for arrays. You can transpose a matrix with `t()`; the generalized equivalent for arrays is `aperm()`. Matrices and arrays can be one dimensional, but will behave differently than vectors. 

## Data Frames

Data frames are lists of equal length vectors. It shares the properties of boths lists and matrices. 

```{r}
## You create a data frame using data.frame(), which takes named vectors as input
df <- data.frame(x = 1:3, y = c("a", "b", "c"))
str(df)

## data.frame will turn strings into factors by default, so be aware
```

You can also coerce an object into a data frame using `ad.data.frame()`. 

  * A vector will create a one-column data frame.

  * A list will create one column for each element; it’s an error if they’re not all the same length.

  * A matrix will create a data frame with the same number of columns and rows as the matrix.
  
You can combine dataframes column-wise with `cbind()` and row-wise with `rbind()`. When combining column-wise, the number of rows must match, but row names are ignored. When combining row-wise, both the number and names of columns must match.

If you want to smoosh a list into a cell of a data frame, use the `I()` function. 
